---
title: "thesis code"
author: "Jaime Ortiz"
date: "7/7/2021"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    theme: united
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare the R environment
Install and load R packages we need for the exercise, and clean the workplace.
```{r, results = 'hide', message = FALSE, warning = FALSE}

## General libraries
library(tidyr)
library(tidyverse)
library(gdata)
library(plyr)
library(dplyr)
library(reshape2)
library(factoextra)
library(lubridate)
library(xlsx)
library(hausekeep)
library(ggbeeswarm)
library(expss)

## Optimal contribution selection
library(data.table)
library(optiSel)

## Plotting libraries
library(ggplot2)
library(ggpubr)
library(UpSetR)
library(remotes)
library(ComplexHeatmap)
rm(list = ls())
```

## Data description (ped)

Pedigree
animal: Identification number animal.
sire: Identification number father. 
dam: Identification number mother. 

geb_dat: Date of birth.
sex: Male or Female.
Basic_Animal:	Identification of base animals.
NBA: EBV - number of born alive (number of piglets).	
DailyGain: EBV - numeric variable. Daily gain (gr/day).	
Backfat: EBV - numeric variable. Back fat (mm)	
sem: Birth semester from 19982 to 20182.	
new_id: Consecutive number to run software EVA (Evolutionary algorithms).	
par20182: Parents of animals born in 20182.	
maxunic: Maximum number of matings
index:	Selection index, each trait was weighted by 1/3. 
cohort:	Number of cohort based on  birth semester (i.e., cohort 1:19982 cohort 41:20182).

```{r}
ped <- readRDS("D:/Masterthesis/R/output/ped_reduced.rds") # Load "ped" previously saved as .rds file 
```

## Figure 2: Plotting the number of animals by cohort

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
# Total number of boars and dams per cohort
tot_count  rbind(count(ped[ped$sex==1,]$cohort),
                   count(ped[ped$sex==2,]$cohort))
tot_count$sex <- rep(c("boar","dam"), each = 41)
tot <- ggplot(tot_count, aes(x=x,y=freq, group=sex))+
  geom_line(aes(color=sex), size = 1)+
  geom_point(aes(color=sex), size = 1.5)+
  scale_color_manual(values=c("yellowgreen", "blue"))+
  labs(x="cohort", y = "count")+
  theme_bw()+
  scale_x_continuous(breaks = seq(1,41,5))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="top") + 
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

# Total number of founder animals per cohort
unknown <- ped[ped$sire == 0 & ped$dam == 0,]
unknown <- as.data.frame(
  aggregate(unknown$animal, list(unknown$cohort,unknown$sex),
            FUN = length)
)
unknown$Group.2 <- ifelse(unknown$Group.2 == 1, "boar", "dam")
colnames(unknown) <- c("cohort", "sex", "count")
un_plot <- ggplot(unknown, aes(x = cohort, y = count, group = sex))+
  geom_line(aes(color = sex), size = 1)+
  geom_point(aes(color = sex), size = 1.5)+
  scale_color_manual(values = c("yellowgreen", "blue"))+
  labs(x = "cohort", y = "count")+
  theme_bw()+
  scale_x_continuous(breaks = seq(1,36,5))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="top") + 
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

ggpubr::ggarrange(tot,un_plot,
                  labels = c("a","b"),heights = c(1, 1), ncol = 1, nrow = 2,
                  common.legend = TRUE, legend = "bottom")
#ggsave(filename = "D:/Masterthesis/graphs/tot_found.png",width=10,height=2*3,dpi=300)
```

## Table 5 and Figure 3: Quality of the pedigree

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
summary_eva_files <- list.files(path = "D:/Masterthesis/R/input/PCI",
                                pattern = "F_summary.txt$",
                                all.files = TRUE, full.names = TRUE)
list_names <- c("gen1","gen2","gen3","gen4","gen5","gen6")
list_eva_summary <- list()
for (i in 1:length(summary_eva_files)) {
  list_eva_summary[[i]] <- read.table(summary_eva_files[i], header = T)
}
names(list_eva_summary) <- list_names

for (i in 1:length(list_eva_summary)) {
  colnames(list_eva_summary[[i]])[colnames(list_eva_summary[[i]]) == 
                                    names(list_eva_summary[[i]])] <-
    c("cohort","n","ninbreed",
      "avgF","maxF","avgcoa",
      "expF","alpha","PCI","gencoef",
      "genint")
}
summary_gen <- do.call(rbind.data.frame,list_eva_summary) # join all data base from eva output
summary_gen$sem <- seq(1:41)
summary_gen$depth <- rep(c("PCI1","PCI2","PCI3","PCI4","PCI5","PCI6"),each = 41)
summary_gen$cohort <- 1:nrow(summary_gen)

## Correlation PCI - Eq
round(cor(summary_gen[,c(3:8)],summary_gen$gen1.Gencoef),digits = 2)

## PCI by generation six generation
PCI <- ggplot(summary_gen, aes(x = sem, y = PCI, group = depth))+
  geom_line(aes(color = depth))+
  labs(x = "Cohort", y = "PCI")+
  theme_bw()+
  scale_x_continuous(breaks = seq(1,41,2))+
  scale_y_continuous(breaks = seq(0,1,.2))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="top") + 
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

## Equi - gen
summary_gen[summary_gen$depth == "PCI1",]$sem
EQ <- ggplot(summary_gen[summary_gen$depth == "PCI1",],
             aes(x = summary_gen[summary_gen$depth == "PCI1",]$sem, 
                 y = summary_gen[summary_gen$depth == "PCI1",]$gencoef))+
  geom_line(size = 1, color = "yellowgreen")+
  labs(x = "Cohort", y = "EqGi")+
  theme_bw()+
  scale_x_continuous(breaks = seq(1,41,2))+
  scale_y_continuous(breaks = seq(0,15,2))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position = "top") + 
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

ggarrange(EQ,PCI, labels = c("a","b"), ncol = 1, nrow = 2,
          common.legend = TRUE, legend = "bottom")
#ggsave(filename = "D:/Masterthesis/graphs/PCI_Eq.png",width=10,height=2*3,dpi=300)
```

## Figure 4: Generation time

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
# boars_age
ped$age_b <- NA
for (i in 1:nrow(ped)) {
  if (ped[i,"sire"] == 0){
    ped[i,"age_b"] = 0
  } else {
    ped[i,"age_b"] = as.numeric(ped[i,"geb_dat"]-
                                  ped[ped$animal == ped[i,"sire"],"geb_dat"])/365
  }
}

# sow_age
ped$age_s <- NA
for (i in 1:nrow(ped)) {
  if (ped[i,"dam"] == 0){
    ped[i,"age_s"] = 0
  } else {
    ped[i,"age_s"] = as.numeric(ped[i,"geb_dat"]-
                                  ped[ped$animal == ped[i,"dam"],"geb_dat"])/365
  }
}

Tg_boar <- aggregate(ped$age_b, list(ped$cohort), FUN = "mean", na.rm = TRUE)
Tg_dam <- aggregate(ped$age_s, list(ped$cohort), FUN = "mean", na.rm = TRUE)    
Tg <- rbind(Tg_boar[Tg_boar$Group.1 > 6, ],Tg_dam[Tg_dam$Group.1 > 6, ])
Tg$sex <- rep(c("boar","dam"), each = 35)
colnames(Tg) <- c("cohort","Tg","sex")

ggplot(Tg, aes(x = cohort, y = Tg, group = sex))+
  geom_line(aes(color = sex), size = 1)+
  geom_point(aes(color = sex), size = 1.5)+
  scale_color_manual(values = c("yellowgreen", "blue"))+
  labs(x = "cohort", y = "Tg")+
  theme_bw()+
  scale_x_continuous(breaks = seq(7,42,3))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="bottom") + 
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

#ggsave(filename = "D:/Masterthesis/graphs/tg.png",width=10,height=2*3,dpi=300)
```

## Figure 5: Average inbreeding, average coancestry and ð¹ð¼ð‘†

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
evaout <- read.table("D:/Masterthesis/R/input/summary.txt", header = T)
evaout$fix <- evaout$avgcoa-1/2/evaout$N # correction of animal coancestry f/2N
evaout$Cohort <- 1:nrow(evaout)

inbre_coa <- evaout[,c(4,8,12)]
colnames(inbre_coa) <- c("Average inbreeding","FIS", "Average coancestry")
summary(inbre_coa)
inbre_coa <- melt(inbre_coa)
inbre_coa$cohort <- rep(c(1:41),3)

ggplot(inbre_coa, aes(x = cohort, y = value, group = variable))+
  geom_line(aes(color = variable), size = 1)+
  geom_point(aes(color = variable), size = 1)+
  scale_color_manual(values = c("yellowgreen", "blue", "orange"))+
  labs(x = "cohort", y = " ")+
  theme_bw()+
  scale_x_continuous(breaks = seq(1,41,3))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="bottom") + 
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

#ggsave(filename = "D:/Masterthesis/graphs/inbreeding_coancestry.png", width=10,height=1*3,dpi=300)
```

## Figure 6: Overall changes in ð¹ð‘†ð‘‡

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
df <- data.frame(x1=8,x2=23,y1=0.0001,y2=0.0134,x3=24,x4=41,y3=0.0071,y4=0.0532)
Fst <- ggplot(evaout, aes(x=Cohort, y=expF))+
  geom_point(size = 2, pch = 16)+
  labs(x = "Cohort", y = "drift coefficient (Fst)")+
  theme_bw()+
  scale_x_continuous(breaks = seq(1,41,2))+
  scale_y_continuous(breaks = seq(0,.05,.01))+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_text(size = 16),
    axis.title.x = element_text(size = 16),
    text = element_text(size = 16))+
  theme(
    axis.line = element_line(colour = "gray20"), 
    panel.grid.major = element_line(colour = "gray85"), 
    panel.grid.minor = element_line(colour = NA))
Fst + 
  geom_segment(aes(x = x1, y = y1,xend = x2, yend = y2,colour = "segment"),
               data = df, size = 1.5, color="yellowgreen")+
  geom_segment(aes(x = x3, y = y3, xend = x4, yend = y4, colour = "segment"),
               data = df, size = 1.5, color="blue")
#ggsave(filename = "D:/Masterthesis/graphs/drift_coeff.png",width=10,height=2*3,dpi=300)

```

## Effective population size

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
# average generation time Line_1 and Line_2
line_1_year <- c(20021,20022,20031,20032,20041,
                 20042,20051,20052,20061,20062,
                 20071,20072,20081,20082,20091,
                 20092) 
line_2_year <- c(20101,20102,20111,20112,20121,
                 20122,20131,20132,20141,20142,
                 20151,20152,20161,20162,20171,
                 20172,20181,20182)

boar_line_1 <- mean(ped[ped$sem%in%line_1_year,]$age_b,na.rm = TRUE)
saw_line_1 <- mean(ped[ped$sem%in%line_1_year,]$age_s,na.rm = TRUE)
(boar_line_1+saw_line_1)/2

boar_line_2 <- mean(ped[ped$sem%in%line_2_year,]$age_b,na.rm = TRUE)
saw_line_2 <- mean(ped[ped$sem%in%line_2_year,]$age_s,na.rm = TRUE)
(boar_line_2+saw_line_2)/2

line_1 <- evaout[8:23,]
summary(lm(expF ~ Cohort, data = line_1))
1/((2*0.001061)*((boar_line_1+saw_line_1)/2))
line_2 <- evaout[24:41,]
summary(lm(expF ~ Cohort, data = line_2))
1/((2*0.002642)*((boar_line_2+saw_line_2)/2))

```

## Table 6 and Figure 7: Plotting genetic trend

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
ped2 = aggregate(ped[,c(7,8,9)], by = list(ped$cohort), FUN = mean, na.rm = TRUE)
ped2[ped2$Group.1 == c(1,23,24,41),]

ebv <- function(data,x,cut){
  ebv_summary <- seWithin(data = data,
                          measurevar = c("trait"),
                          withinvars = c("cohort"),
                          idvar = "cohort")
  plt <- ggplot(ebv_summary, aes(cohort,trait)) +
    geom_quasirandom(data = data, alpha = 0.1, colour = "yellowgreen") + # this line plots raw data and can be omitted, depending on your plotting
    geom_point(colour = "orange") + # add point
    geom_errorbar(aes(ymin = trait - ci, ymax = trait + ci),
                  width = 0, colour = "blue") +
    labs(x = "Cohort", y = x) +
    theme_bw()+
    scale_x_continuous(breaks = seq(1,41,1)) +
    scale_y_continuous(breaks = cut) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
    theme(legend.position="top")+
    theme(axis.line = element_line(colour = "gray20"),
          panel.grid.major = element_line(colour = "gray85"),
          panel.grid.minor = element_line(colour = NA))
  return(plt)
}

# NBA
ped_trait <- ped[,c(15,7)]
colnames(ped_trait) <- c("cohort","trait")
summary(ped_trait)
x <- "NBA(units)"
cut <- seq(-2,3.6,1)
plt_nba <- ebv(ped_trait, x, cut)
plt_nba + 
  geom_segment(aes(x = 1, y = -0.01,xend = 23, yend = -0.22,colour = "segment"),
               data = df, size = 1, color="blue")+
  geom_segment(aes(x = 24, y = 0.13, xend = 41, yend = 1.97, colour = "segment"),
               data = df, size = 1, color="blue")

# Linear regression NBA before and after cohort 23
summary(lm(NBA ~ cohort, data = ped[ped$cohort <= 23,]))
summary(lm(NBA ~ cohort, data = ped[ped$cohort > 23,]))

# Daily Gain
ped_trait <- ped[,c(15,8)]
colnames(ped_trait) <- c("cohort","trait")
summary(ped_trait)
x <- "Daily gain (gr/day)"
cut <- seq(-208,206,100)
plt_dg <- ebv(ped_trait, x, cut)
plt_dg +
  geom_segment(aes(x = 1, y = -1.26,xend = 23, yend = -18.4,colour = "segment"),
               data = df, size = 1, color="blue")+
  geom_segment(aes(x = 24, y = 10.31, xend = 41, yend = 63.46, colour = "segment"),
               data = df, size = 1, color="blue")

# Linear regression daily gain before and after cohort 23
summary(lm(DailyGain ~ cohort, data = ped[ped$cohort <= 23,]))
summary(lm(DailyGain ~ cohort, data = ped[ped$cohort > 23,]))

# Back fat thickness
ped_trait <- ped[,c(15,9)]
colnames(ped_trait) <- c("cohort","trait")
summary(ped_trait)
x <- "Back fat (mm)"
cut <- seq(-3,5,1)
plt_bf <- ebv(ped_trait, x, cut)
plt_bf +
  geom_segment(aes(x = 1, y = 0.03,xend = 23, yend = 0.34,colour = "segment"),
               data = df, size = 1.5, color="blue")+
  geom_segment(aes(x = 24, y = 0.31, xend = 41, yend = 0.26, colour = "segment"),
               data = df, size = 1.5, color="blue")

# Linear regression Back fat thickness before and after cohort 23
summary(lm(Backfat ~ cohort, data = ped[ped$cohort <= 23,]))
summary(lm(Backfat ~ cohort, data = ped[ped$cohort > 23,]))

#ggpubr::ggarrange(plt_nba,plt_dg,plt_bf, labels = c("a","b","c"),heights = c(1, 1), ncol = 1, nrow = 3, common.legend = TRUE, legend = "bottom")
```

## Figure 8: Scaling original traits

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
# Filtering animals born in the second half of 2018
born20182 <- ped[ped$sem == 20182,]
# Scaling original values
ped$NBA_scal <- (ped$NBA)/sd(ped[ped$animal%in%unique(c(born20182$sire,born20182$dam)),]$NBA)
ped$Daygain_scal <- (ped$DailyGain)/sd(ped[ped$animal%in%unique(c(born20182$sire,born20182$dam)),]$DailyGain)
ped$Backfat_scal <- (ped$Backfat)/sd(ped[ped$animal%in%unique(c(born20182$sire,born20182$dam)),]$Backfat)

# values before scaling
ped_before_scaling <- ped[,c(1,7,8,9)]
colnames(ped_before_scaling) <- c("animal","NBA","Daily Gain","Backfat")  
ped_before_scaling <- melt(ped_before_scaling, id = "animal")

ped_after_scaling <- ped[,c(1,16,17,18)]
colnames(ped_after_scaling) <- c("animal","NBA","Daily Gain","Backfat")  
ped_after_scaling <- melt(ped_after_scaling, id = "animal")

plot1 <- ped_before_scaling %>%
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(width = 0.8) + coord_flip() +
  scale_x_discrete(expand = c(2.5,0)) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3,
               color = "red", fill = "red") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.title.y = element_text(size = 15),
          axis.title.x = element_text(size = 15),
          text = element_text(size = 15)) + 
    theme(panel.grid.major = element_line(colour = "gray91"),
          panel.background = element_rect(fill = NA)) + 
    theme(plot.background = element_rect(colour = "black",linetype = "solid"))+
    labs(x = NULL)

plot2 <- ped_after_scaling %>%
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(width = 0.8) + coord_flip() +
  scale_x_discrete(expand = c(2.5,0)) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3,
               color = "red", fill = "red") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15)) + 
  theme(panel.grid.major = element_line(colour = "gray91"),
        panel.background = element_rect(fill = NA)) + 
  theme(plot.background = element_rect(colour = "black",linetype = "solid")) +
  labs(x = NULL)

ggpubr::ggarrange(plot1,plot2,
                  labels = c("a","b"),heights = c(1, 1), ncol = 1, nrow = 2,
                  common.legend = TRUE, legend = "bottom")
ggsave(filename = "D:/Masterthesis/graphs/before_after_scaling1.png", width=13,height=2*3,dpi=300)

```

## Table 7: Uncertainty of the model

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
uncer <- function(ex){
  ex_list <- list()
  for (i in 1:length(ex)) {
    ex_list[[i]] <- read.table(ex[i], header = T)
  }
  names(ex_list) <- c("rep1","rep2","rep3","rep4","rep5")
  ## ----- create data.frame to storage the average values
  uncer <- data.frame(
    measurements <- c("rep1","rep2","rep3","rep4","rep5"),
    matrix(data = NA, ncol = 2, nrow = 5)
  )
  colnames(uncer) <- c("measurements","EBV","F")
  ## ----- filling up uncertainty data.frame
  # genetic merit
  uncer[1,"EBV"] <- mean(ex_list[[1]]$EBV)
  uncer[2,"EBV"] <- mean(ex_list[[2]]$EBV)
  uncer[3,"EBV"] <- mean(ex_list[[3]]$EBV)
  uncer[4,"EBV"] <- mean(ex_list[[4]]$EBV)
  uncer[5,"EBV"] <- mean(ex_list[[5]]$EBV)
  # expected inbreeding 
  uncer[1,"F"] <- mean(ex_list[[1]]$F)
  uncer[2,"F"] <- mean(ex_list[[2]]$F)
  uncer[3,"F"] <- mean(ex_list[[3]]$F)
  uncer[4,"F"] <- mean(ex_list[[4]]$F)
  uncer[5,"F"] <- mean(ex_list[[5]]$F)
  
  print(uncer)
  
  print(sd(uncer$EBV)/sqrt(nrow(uncer))) # standard error genetic merit
  print(sd(uncer$F)/sqrt(nrow(uncer))) # standard error expected inbreeding offspring
}

ex1 <- list.files(c("D:/Masterthesis/eva/ocs_index/ex1/rep1",
                    "D:/Masterthesis/eva/ocs_index/ex1/rep2",
                    "D:/Masterthesis/eva/ocs_index/ex1/rep3",
                    "D:/Masterthesis/eva/ocs_index/ex1/rep4",
                    "D:/Masterthesis/eva/ocs_index/ex1/rep5"),
                  pattern = "MatingList.txt$",
                  all.files = TRUE, full.names = TRUE ) 
uncer(ex1)
```

## Table 8: Selected boars per mating and per scenario from 40 boars candidates

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
candidates <- list.files(c("D:/Masterthesis/eva/ocs_index/ex1/rep3",
                           "D:/Masterthesis/eva/ocs_index/ex2/rep3",
                           "D:/Masterthesis/eva/ocs_index/ex3/rep3",
                           "D:/Masterthesis/eva/ocs_index/ex4/rep3",
                           "D:/Masterthesis/eva/ocs_index/ex5",
                           "D:/Masterthesis/eva/ocs_index/ex6/rep3",
                           "D:/Masterthesis/eva/ocs_index/ex7/rep3"),
                           pattern = "candidates.txt$",
                         all.files = TRUE,
                         full.names = TRUE ) 

list_names <- c("run1","run2","run3","run4","run5","run6","run7")
candidates_list <- list()
for (i in 1:length(candidates)) {
  candidates_list[[i]] <- read.table(candidates[i], header = T)
}

new_list <- list()
for (i in 1:7) {
  new_list[[i]] <- candidates_list[[i]][candidates_list[[i]]$Sex == 1 & 
                                         candidates_list[[i]]$N.matings != 0,]  
}

cand <- 
  left_join(new_list[[1]][,c(1,3)], new_list[[2]][,c(1,3)], by="ID") %>% 
  left_join(., new_list[[3]][,c(1,3)], by="ID") %>%
  left_join(., new_list[[4]][,c(1,3)], by="ID") %>%
  left_join(., new_list[[5]][,c(1,3)], by="ID") %>%
  left_join(., new_list[[6]][,c(1,3)], by="ID") %>%
  left_join(., new_list[[7]][,c(1,3)], by="ID") 

colnames(cand) <- c("id","uno","dos","tres","cuatro","cinco","seis","siete")
cand$index <- ped[ped$animal %in% unique(cand$id),]$index

head(cand)

write.xlsx(cand, file = "D:/Masterthesis/graphs/cand.xlsx", sheetName="cand", append=TRUE)

matings_output<-list.files(c("D:/Masterthesis/eva/ocs_index/ex1/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex2/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex3/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex4/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex5",
                             "D:/Masterthesis/eva/ocs_index/ex6/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex7/rep3"),
                           pattern = "MatingList.txt$",all.files = TRUE,
                           full.names = TRUE) 
list_names <- c("run1","run2","run3","run4","run5","run6","run7")
ocs_output_matings <- list()
for (i in 1:length(matings_output)) {
  ocs_output_matings[[i]] <- read.table(matings_output[i], header = T)
}
names(ocs_output_matings) <- list_names

scenario <- c("(1,0)","(1,-100)","(1,-300)",
              "(1,-1000)","(1,-3000)","(1,-10000)","(0,-1)")
number <- c(length(unique(ocs_output_matings[[1]]$Sire)),
            length(unique(ocs_output_matings[[2]]$Sire)),
            length(unique(ocs_output_matings[[3]]$Sire)),
            length(unique(ocs_output_matings[[4]]$Sire)),
            length(unique(ocs_output_matings[[5]]$Sire)),
            length(unique(ocs_output_matings[[6]]$Sire)),
            length(unique(ocs_output_matings[[7]]$Sire)))
tibble(scenario,number)
```

## Upset plot function

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}

comparison <- function(list,x){
  # plot general mating
  m <- make_comb_mat(
    list_to_matrix(
      list("1,0"=list[[1]],
           "1,-100"=list[[2]],
           "1,-300"=list[[3]],
           "1,-1000"=list[[4]],
           "1,-3000"=list[[5]],
           "1,-10000"=list[[6]],
           "0,-1"=list[[7]])
    )
  )
  ss = set_size(m)
  cs = comb_size(m)
  ht = UpSet(m, 
             set_order = c("1,0","1,-100",
                           "1,-300","1,-1000",
                           "1,-3000","1,-10000",
                           "0,-1"),
             #comb_order = order(comb_degree(m),-cs),
             top_annotation = HeatmapAnnotation(
               "Intersections" = anno_barplot(cs,
                                              ylim = c(0, max(cs)*1.1),
                                              border = FALSE, 
                                              gp = gpar(fill = "yellowgreen"), 
                                              height = unit(2, "cm")
               ), 
               annotation_name_side = "left", 
               annotation_name_rot = 90),
             left_annotation = rowAnnotation(
               "Pop.size" = anno_barplot(-ss,
                                         baseline = 0,
                                         axis_param = list(
                                           at = x*-1,
                                           labels = x,
                                           labels_rot = 0),
                                         border = FALSE,
                                         gp = gpar(fill = "yellowgreen"), 
                                         width = unit(4, "cm")
               ),
               set_name = anno_text(set_name(m), 
                                    location = 1, 
                                    just = "right",
                                    width = max_text_width(set_name(m)) + 
                                      unit(4, "mm"))
             ), 
             right_annotation = NULL,
             show_row_names = FALSE)
  ht = draw(ht)
  od = column_order(ht)
  decorate_annotation("Intersections", {
    grid.text(cs[od], 
              x = seq_along(cs),
              y = unit(cs[od], "native") + unit(6, "pt"), 
              default.units = "native", just = c("center", "bottom"), 
              gp = gpar(fontsize = 9, col = "#404040"), rot = 0)
  })
}
```

Figure 9: Upset plot - selected boars

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
matings_output<-list.files(c("D:/Masterthesis/eva/ocs_index/ex1/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex2/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex3/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex4/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex5",
                             "D:/Masterthesis/eva/ocs_index/ex6/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex7/rep3"),
                           pattern = "MatingList.txt$",all.files = TRUE,
                           full.names = TRUE) 
unique_male <- list()
for (i in 1:length(matings_output)){
  unique_male[[i]] <- unique(ocs_output_matings[[i]]$Sire)
}

#png("D:/Masterthesis/graphs/compa_index_male.png",
#    width = 40, height = 20, units = 'cm', res = 300)
y <- c(0,10,20,30,40)
comparison(unique_male,y)
#dev.off()

```

Figure 10: Upset plot complete mating (boar*dam)

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
matings_output<-list.files(c("D:/Masterthesis/eva/ocs_index/ex1/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex2/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex3/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex4/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex5",
                             "D:/Masterthesis/eva/ocs_index/ex6/rep3",
                             "D:/Masterthesis/eva/ocs_index/ex7/rep3"),
                           pattern = "MatingList.txt$",all.files = TRUE,
                           full.names = TRUE) 
list_names <- c('run1','run2','run3','run4','run5','run6','run7')
ocs_output_matings <- list()
for (i in 1:length(matings_output)){
  ocs_output_matings[[i]] <- read.table(matings_output[i], header = T)
}
names(ocs_output_matings) <- list_names
# adding mating column to compare mating between run when the index was used as a selection criteria
for (i in 1:length(matings_output)){
  ocs_output_matings[[i]]$mating <- paste0(ocs_output_matings[[i]]$Sire,
                                           ocs_output_matings[[i]]$Dam)
}
unique_list <- list()
for (i in 1:length(matings_output)){
  unique_list[[i]] <- unique(ocs_output_matings[[i]]$mating)
}

#png("D:/Masterthesis/graphs/compa_index_all.png",
#    width = 40, height = 20, units = 'cm', res = 300)
x <- c(0,200,397)
comparison(unique_list,x)
#dev.off()

```

Figure 11: Selection differential

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
eff.diff <- function(eva_mat, ped , x = 1, y = 2, sem = 20182){
  diff      <- as.data.frame(table(eva_mat[x]))
  index     <- ped[ped$animal %in% diff$Var1,]$index
  NBA       <- ped[ped$animal %in% diff$Var1,]$NBA
  DailyGain <- ped[ped$animal %in% diff$Var1,]$DailyGain
  Backfat   <- ped[ped$animal %in% diff$Var1,]$Backfat
  diff <- cbind(diff,index,NBA,DailyGain,Backfat)
  colnames(diff) <- c("id","count","index","NBA","DailyGain","Backfat")
  
  cand <- unique(ped[ped$sem == sem,][[y]])
  
  eff.diff <- data.frame(
    trait <- c("index","NBA","DailyGain","Backfat"),
    average <- c(
      weighted.mean(diff$index,diff$count),
      weighted.mean(diff$NBA,diff$count),
      weighted.mean(diff$DailyGain,diff$count),
      weighted.mean(diff$Backfat,diff$count)
    ),
    eff.diff <- c(
      weighted.mean(diff$index,diff$count)- 
        mean(ped[ped$animal %in% cand,]$index),
      weighted.mean(diff$NBA,diff$count)- 
        mean(ped[ped$animal %in% cand,]$NBA),
      weighted.mean(diff$DailyGain,diff$count)- 
        mean(ped[ped$animal %in% cand,]$DailyGain),
      weighted.mean(diff$Backfat,diff$count)- 
        mean(ped[ped$animal %in% cand,]$Backfat)
    )
  )
  colnames(eff.diff) <- c("trait","avg.sel","diff.sel")
  eff.diff
}

diff.frame <- rbind(
  eff.diff(ocs_output_matings[[1]], ped, x = 1, y = 2, sem = 20182),
  eff.diff(ocs_output_matings[[2]], ped, x = 1, y = 2, sem = 20182),
  eff.diff(ocs_output_matings[[3]], ped, x = 1, y = 2, sem = 20182),
  eff.diff(ocs_output_matings[[4]], ped, x = 1, y = 2, sem = 20182),
  eff.diff(ocs_output_matings[[5]], ped, x = 1, y = 2, sem = 20182),
  eff.diff(ocs_output_matings[[6]], ped, x = 1, y = 2, sem = 20182),
  eff.diff(ocs_output_matings[[7]], ped, x = 1, y = 2, sem = 20182)
)

diff.frame$scenario <- rep(c('1,0','1,-100','1,-300','1,-1000',
                             '1,-3000','1,-10000','0,-1'), each = 4)

ggplot(
  data = diff.frame[diff.frame$trait == "index" &
                      diff.frame$scenario != "1,0",],
  aes(x = avg.sel, 
      y = diff.sel))+
  geom_point(size = 10, aes(color = scenario))+
  geom_hline(yintercept = 0.62, color='blue',linetype='dotted')+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="bottom") + 
  labs(x = "Average on Index for selected boars by scenario",
       y = "Differential Selection (S)")+
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

#ggsave(
#  filename=
#    "D:/Masterthesis/graphs/diff.sel_index.png",
#  width=15,height=2*3,dpi=300)
```

Figure 12: Rate of inbreeding and change in average coancestry

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
evacoeff <- read.table("D:/Masterthesis/R/input/evacoeff.txt", header = T) # this is an outcome from EVA to make quality graphs

delta <- function(list,evacoeff){
  #  Create a list of mating from EVA
  ocs_output_matings <- list()
  for (i in 1:length(list)) {
    ocs_output_matings[[i]] <- read.table(list[i], header = T)
  }
  # Making a data.frame
  delta <- data.frame(
    run <- c("1,0","1,-100","1,-300","1,-1000","1,-3000","1,-10000","0,-1"),
    matrix(data = NA, ncol = 2, nrow = length(ocs_output_matings))
  )
  names(delta) <- c("run","rate of inbreeding","change avg.coancestry")
  # Computing expected inbreeding for the next cohort
  for (i in 1:length(ocs_output_matings)) {
    # Computing rate of inbreeding
    delta[i,2] <- 
      ((mean(ocs_output_matings[[i]]$F)-
          mean(evacoeff[evacoeff$Animal %in% 
                          c(unique((ocs_output_matings[[i]])$Sire),
                            (ocs_output_matings[[i]]$Dam)),]$F))/
         (1-mean(evacoeff[evacoeff$Animal %in% 
                            c(unique((ocs_output_matings[[i]])$Sire),
                              (ocs_output_matings[[i]]$Dam)),]$F)))*100 
    # Computing change in average coancestry
    A <- makeA(ped,keep.only = c(unique(ocs_output_matings[[i]]$Sire),
                                 unique(ocs_output_matings[[i]]$Dam)))
    c_frame <- data.frame(table(c(ocs_output_matings[[i]]$Sire,
                                  ocs_output_matings[[i]]$Dam)))
    c <- c_frame$Freq/sum(c_frame$Freq)
    delta[i,3] <- ((c%*%A%*%c-mean(A))/(1-mean(A)))*100
  }

  delta <- reshape2::melt(delta)
  delta
}

list <- list.files(c("D:/Masterthesis/eva/ocs_index/ex1/rep3",
                     "D:/Masterthesis/eva/ocs_index/ex2/rep3",
                     "D:/Masterthesis/eva/ocs_index/ex3/rep3",
                     "D:/Masterthesis/eva/ocs_index/ex4/rep3",
                     "D:/Masterthesis/eva/ocs_index/ex5",
                     "D:/Masterthesis/eva/ocs_index/ex6/rep3",
                     "D:/Masterthesis/eva/ocs_index/ex7/rep3"),
                   pattern = "MatingList.txt$",
                   all.files = TRUE,
                   full.names = TRUE)


ocs_index <- delta(list,evacoeff)

p1 <- ggplot(data = ocs_index, aes(x=run, y = value, fill = variable))+
  geom_bar(stat = "identity", position=position_dodge()) +
  labs(x = "scenario", y = "%")+
  theme_bw()+
  scale_y_continuous(breaks = seq(-1,2,.5))+
  scale_x_discrete(limits = c("1,0","1,-100",
                              "1,-300","1,-1000",
                              "1,-3000","1,-10000",
                              "0,-1"))+
  scale_fill_manual(name = " ",
                    values = c("yellowgreen","blue"))+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 12, angle = 90),
        axis.title.x = element_text(size = 12),
        text = element_text(size = 15))+
  theme(legend.position="bottom") +
  theme(axis.line = element_line(colour = "gray20"), 
    panel.grid.major = element_line(colour = "gray85"), 
    panel.grid.minor = element_line(colour = NA))
p1

ggsave(filename='D:/Masterthesis/graphs/delta_rel_index_full.png', width=10, height=2*3, dpi=300)
```

Figure 13: Differential selection per average coancestry

```{r, results="hide", fig.show="hide", message = FALSE, warning = FALSE}
scenario <- c('1,0','1,-100','1,-300','1,-1000','1,-3000','1,-10000','0,-1')
coa <- c(1.68,1.68,1.65,1.63,1.27,0.82,0.85)
diff <- c(0.61,0.61,0.61,0.61,0.49,0.11,-0.014)
coa_diffsel <- data.frame(scenario,coa,diff)
coa_diffsel$coa2 <- coa_diffsel$coa^2

ggplot(
  data = coa_diffsel[coa_diffsel$scenario != "1,0",],
  aes(x = coa,
      y = diff,
      fill = scenario)) +
  geom_point(size = 10, aes(color = scenario))+
  geom_hline(yintercept = 0.61, color='blue',linetype='dotted')+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        text = element_text(size = 15))+
  theme(legend.position="bottom") + 
  labs(x = "Change in average coancestry",
       y = "Differential Selection (S)")+
  theme(axis.line = element_line(colour = "gray20"),
        panel.grid.major = element_line(colour = "gray85"),
        panel.grid.minor = element_line(colour = NA))

ggsave(filename="D:/Masterthesis/graphs/diff.sel.per.coanc.png", width=15,height=2*3,dpi=300)

## Quadratic model

scenario <- c('1,0','1,-100','1,-300','1,-1000','1,-3000','1,-10000','0,-1')
coa <- c(1.68,1.68,1.65,1.63,1.27,0.82,0.85)
diff <- c(0.61,0.61,0.61,0.61,0.49,0.11,-0.014)
coa_diffsel <- data.frame(scenario,coa,diff)
coa_diffsel$coa2 <- coa_diffsel$coa^2

summary(lm(coa_diffsel[coa_diffsel$scenario != "0,-1",]$diff ~ 
             coa_diffsel[coa_diffsel$scenario != "0,-1",]$coa +
             coa_diffsel[coa_diffsel$scenario != "0,-1",]$coa2
))
```
